Assessing the impacts of time to detection distribution assumptions on detection probability estimation

### 2017-04-18 ReviewerComments.tex
Add my comments on reviewer response

I have included several items for the authors to consider below and provide this review as an Ecologist and potential model user.
Several pieces of information are lacking from the manuscript that will be relevant to users of this method.
\begin{enumerate}
\item What types of data are relevant to each examined distribution (e.g., gamma, Weibull)?
\item What sample sizes (i.e., number of detections) are needed to obtain reliable estimates with CIs that are not so wide as to be meaningless?
\item What are approximate computation times for models of different complexity (e.g., days? weeks?).
\end{enumerate}
\begin{quote}
We address each of these in turn as they appear in the comments below.
\end{quote}

@@

Line 40-41.  Also worth mentioning caveats - e.g., greater imprecision and computation times.
\begin{quote}
\adam{I certainly don't intend to address either of these in the Abstract.  Both are discussed in the text.  Computation times are not addressed in the Discussion.  Let me know if you think otherwise.}
\end{quote}

Introduction\\
Line 36-41: I find mention of survival analyses here and later in the manuscript confusing for most readers - especially Ecologists, presuming this is the intended audience.  It may be due to the unorthodox use of time-to-detection instead of traditional time-to-event analyses often used to estimate survival.  I suggest either omitting the reference to survival analysis here and elsewhere in the manuscript, or using the time-to-event terminology in this sentence and providing citations for the cdf and pdfs outlined here for reference.  Then explain that within the manuscript you will subsequently refer to it as TTDD.  
\begin{quote}
We have made the equivalence more explicit: ``We analyze times to first detection as time-to-event data, as is done in parametric survival analysis, ..."
\end{quote}

@@

We appreciate the importance of providing biological justification for statistical assumptions.
It is always a good foundation.
That being said, the use of any time-to-detection distribution is a statistical approximation to empirical patterns.
If the dynamics of the system are more complicated than we can directly model, the best we can do is find a TTDD that reasonably describes the observed pattern.
For instance, if the arrival of the observer causes non-constant detection, we know neither how long that observer effects lasts nor how the degree of that effect changes over time.
Gamma, lognormal, and Weibull TTDDs all provide different models for that detection pattern with different inferences for tail probability.
\end{quote}

@@

@@

\item Simulations incorporating variable sample sizes to evaluate relative bias and precision would be very useful in terms of study design and applicability.  
\begin{quote}
We have extended our simulation study to include a variety of detection probabilities rather than a variety of sample sizes.
Both affect estimation and are worthy of investigation, but since the performance of unknown-$N$ binomial models generally deteriorates as values of $\pdet$ decrease, we chose to direct our attentions there.
Sections \ref{M-sec:mixture} and \ref{M-sec:family} have been reworked to detail the updated findings.
\end{quote}

\item How might study design impact bias and precision?  For example, length of survey relative to peaked TTDDs and pooling of time periods during data collection?  These are also scenarios that would benefit readers.  Alternatively, these issues could be addressed in the Discussion.
\begin{quote}
These are important decision in the design of surveys, but they are beyond the scope we have defined for this study.
Not having explicitly investigated them, we prefer not to speculate.
%\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
We have separated out a paragraph in the Discussion posing some design-related questions.
\end{quote}

@@

``Computation times for this simulation study were much greater than for the other studies because partitions of the cdf, e.g. $\Fm(C_i|\alpha,\varphi_s)$, had to be calculated separately for every survey; also, sampling often required 2-8 as many iterations.
Average times for exponential, lognormal, Weibull, and gamma model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2, 3, 3.25, and 5.35 minutes for the intercept-only models.
Due to the computation times involved, we fit each model only once."
\end{quote}

\item I don’t fully appreciate the need for both 50\% and 90\% coverage estimates.  It seems one estimate would be sufficient, and typically 90\% or 95\% credible intervals are used with these types of analyses?
\begin{quote}
We agree that using both is redundant.
We now choose to report only 90\% credible intervals.
\end{quote}
\end{enumerate}

Results\\
Page 18, Figure 2:  Holy moly, the CIs of the mixture distributions are exceedingly wide given a sample size of 381 detections (which is a lot relative to many studies).
\begin{quote}
Clarification: We have 947 detections at 381 sites (which is \textit{really} a lot of detections).
We now clearly state in Section \ref{M-sec:data}: ``This dataset includes 947 Ovenbirds counted in 381 surveys at 65 sites with site-specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s."
\end{quote}

@@

As the authors’ suggest, most point-count studies are designed to maximize availability, which is typically high for birds (e.g., 80-95\%) with perceptibility (e.g., the probability an observer detects a cue that is given by an animal) being the much larger source of bias.  Therefore, this model, without including a supplemental method to estimate perceptibility, is of limited use to many readers.
\begin{quote}
We agree.
This manuscript applies to removal-only analyses.
Still, the concept can be carried over to removal-distance analyses, as we describe in a paragraph of the Discussion.
The implementation will be more complicated and will vary depending on how removal- and distance-sampling are integrated.
\end{quote}
Further, although the method may be more accurate even for constant detection data, the loss in precision is such that users will likely not opt to use it unless they suspect non-constant detection through time.  Further, given its presumably arduous computation times (albeit unknown), adding a perceptibility component to the model would likely greatly increase computational demands with unknown implications to bias and precision.  Including a paragraph in the Discussion that outlines when the model would be most beneficial to users (e.g., what type of data, surveys, or sample sizes) and explicitly stating drawbacks (e.g., computation time, loss of precision) would greatly improve this section.  
\begin{quote}
\adam{I do not deem this necessary.  We have not explicitly studied sample size.  The rest is addressed at various locations in the manuscript but does not, I think, require a dedication paragraph in the Discussion.}
\end{quote}

@@

Second, although the original number of simulations was inadequate for characterizing model accuracy, it nonetheless highlighted scenarios of gross inaccuracy, such as: (i) omitting a mixture component in modeling a peaked mixture dataset, or (ii) assuming constant detection when the rate varies during the survey period.
\end{quote}

Second, the statistical properties of the estimator of p(det) are likely depending on the size of the true value of p(det). Therefore, I think Table 1 should have been repeated for a whole range of values of the true value of p(det).
(Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is ``uniformly" good?).
Without variation in ``true" simulated scenarios, it is easy to inadvertendly ``stack the deck" in favor or against a particular combination of settings (Particularly since your ``true" value of p is high: 0.8). A wide range of simulated truths is also necessary because to be useful, these methods should be widely applicable in the
tropics (low N's, low p's very often) as well as in temperate forests (large N's, large p's), for example.
\begin{quote}
We have repeated the simulations in Sections \ref{M-sec:mixture} and \ref{M-sec:family} at true detection probabilities $\pdet = 0.50, 0.65, 0.80$, and 0.95.
The findings have significantly added to our results and conclusions.

With reference to Olkin's finding that point estimates of $N_s$ are unstable unless $\pdet>0.41$, it is not a hard boundary.  
It varies by context, but the boundary is a general property of estimating $N$ in binomial unknown-$N$, unknown-$p$ problems.
Olkin's calculation derives from the relative sample mean and sample variance of repeated large samples with the same values of $N$ and $\pdet$.
As such, it is a best-case scenario.
Similar thresholds continue to recur in various abundance-estimating contexts, as demonstrated in these recent articles: 
\begin{itemize}
\item \citet{Veech2016}: advise caution when $p<0.5$ in the context of individual-level detection effects (no time-to-detection data). 
\item \citet{Field2016}: plots indicate uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.  Methods include removal, double-observer, distance, and multiple-visit.
\item \citet{Davis2016}: report accurate $\hat{N}$ at $\pdet >0.40$ in a traditional removal-sampling context, though for small abundance ($N<50$) they required $\pdet > 0.7$.
\end{itemize}
\end{quote}

@@

We agree that discussions of identifiability and the flatness of the likelihood is very relevant, and we have introduced an explicity discussion of them in the Discussion.

As an informational note, the \texttt{R} package by Solymos was constructed for BUGS and JAGS, though it would not be difficult to adapt data cloning for Stan.
\end{quote}

@@

Writing: The multiplicity of mathematical explicit meanings for the word ``mixture" may result confusing. The distinction between the mixture component a la Farnsworth et al and the N-mixture models (e.g. binomial counts with N~pois and p random) is clear to me, but it may not be for some of JABES' audience.  I suggest re-shaping the introduction to that effect. 
\begin{quote}
In the Introduction, we now clearly refer to a practice of modeling the TTDD ``as a mixture of two distributions --- a continuous-time distribution and a point mass for increased detection probability in the initial observation period."
We define our use of `mixture model' later in the paragraph:  ``We apply the term `mixture model' to any model with a mixture TTDD."
This is contrasted in our definition of `N-mixture model' in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model [\dotso] --- which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
\end{quote}

@@

The authors could have crafted a beautiful and very telling graph (Figure 1) representing the multinomial intervals, the TTDD, the right censoring due to the end of the observation period, and the increased detection probability mixture component. Such figure (which is what I did by hand as I was reading the manuscript) would be a very useful guide in the reading of the paper.
\begin{quote}
We have added this as Figure \ref{M-fig:schematic}.
\end{quote}

>>>>

I have included several items for the authors to consider below and provide this review as an Ecologist and potential model user.
Several pieces of information are lacking from the manuscript that will be relevant to users of this method.
\begin{enumerate}
\item What types of data are relevant to each examined distribution (e.g., gamma, Weibull)?
\item What sample sizes (i.e., number of detections) are needed to obtain reliable estimates with CIs that are not so wide as to be meaningless?
\item What are approximate computation times for models of different complexity (e.g., days? weeks?).
\end{enumerate}
\begin{quote}
We address each of these in turn as they appear in the comments below. 
\jarad{Answer them here even if it repeats information below.}
\end{quote}

@@

Line 40-41.  Also worth mentioning caveats - e.g., greater imprecision and computation times.
\begin{quote}
\adam{I certainly don't intend to address either of these in the Abstract.  Both are discussed in the text.  Computation times are not addressed in the Discussion.  Let me know if you think otherwise.}
\jarad{I think that's fine. In responding to the review, we can state our 
opinion that the abstract is not the appropriate place and point the reviewer
to where the content can be found.} 
\end{quote}

Introduction\\
Line 36-41: I find mention of survival analyses here and later in the manuscript confusing for most readers - especially Ecologists, presuming this is the intended audience.  It may be due to the unorthodox use of time-to-detection instead of traditional time-to-event analyses often used to estimate survival.  I suggest either omitting the reference to survival analysis here and elsewhere in the manuscript, or using the time-to-event terminology in this sentence and providing citations for the cdf and pdfs outlined here for reference.  Then explain that within the manuscript you will subsequently refer to it as TTDD.  
\begin{quote}
We have made the equivalence more explicit: ``We analyze times to first detection as time-to-event data, as is done in parametric survival analysis, ..."
\jarad{We probably need to add more, e.g. intended audience. Philip?} 
\end{quote}

@@

We appreciate the importance of providing biological justification for statistical assumptions.
It is always a good foundation.
That being said, the use of any time-to-detection distribution is a statistical approximation to empirical patterns.
If the dynamics of the system are more complicated than we can directly model, the best we can do is find a TTDD that reasonably describes the observed pattern.
For instance, if the arrival of the observer causes non-constant detection, we know neither how long that observer effects lasts nor how the degree of that effect changes over time.
Gamma, lognormal, and Weibull TTDDs all provide different models for that detection pattern with different inferences detection probability.
\end{quote}

@@

\jarad{We should probably remove TTDD from the introduction.}

@@

\item Simulations incorporating variable sample sizes to evaluate relative bias and precision would be very useful in terms of study design and applicability.  
\begin{quote}
\jarad{I think we can thank reviewers here for the push to do more simulation 
studies that resulted in our realization of a computational savings allowing
us to do more analyses.}
We have extended our simulation study to include a variety of detection probabilities rather than a variety of sample sizes.
Both affect estimation and are worthy of investigation, but since the performance of unknown-$N$ binomial models generally deteriorates as values of $\pdet$ decrease, we chose to direct our attentions there.
\jarad{Don't we have different sample sizes due to the variation in detection
probabilities?}
Sections \ref{M-sec:mixture} and \ref{M-sec:family} have been reworked to detail the updated findings.
\end{quote}

\item How might study design impact bias and precision?  For example, length of survey relative to peaked TTDDs and pooling of time periods during data collection?  These are also scenarios that would benefit readers.  Alternatively, these issues could be addressed in the Discussion.
\begin{quote}
These are important decision in the design of surveys, but they are beyond the scope we have defined for this study.
Not having explicitly investigated them, we prefer not to speculate.
%\citet{Petit1995, Johnson2008, LeeMarsden2008} and \citet{Reidy2011} all address the effects of duration.
We have separated out a paragraph in the Discussion posing some design-related questions.
\jarad{Be more specific.}
\end{quote}

@@

``Computation times for this simulation study were much greater than for the other studies because partitions of the cdf, e.g. $\Fm(C_i|\alpha,\varphi_s)$, had to be calculated separately for every survey; also, sampling often required 2-8 as many iterations.
Average times for exponential, lognormal, Weibull, and gamma model fits were 2.6, 5.1, 5.3, and 25+ hours, respectively, as compared to only 2, 3, 3.25, and 5.35 minutes for the intercept-only models.
\jarad{Too many significant digits.}
Due to the computation times involved, we fit each model only once."
\end{quote}

\item I don’t fully appreciate the need for both 50\% and 90\% coverage estimates.  It seems one estimate would be sufficient, and typically 90\% or 95\% credible intervals are used with these types of analyses?
\begin{quote}
We agree that using both is redundant.
We now choose to report only 90\% credible intervals.
\jarad{Are 50\% included in supplement?}
\end{quote}
\end{enumerate}

Results\\
Page 18, Figure 2:  Holy moly, the CIs of the mixture distributions are exceedingly wide given a sample size of 381 detections (which is a lot relative to many studies).
\begin{quote}
Clarification: We have 947 detections at 381 sites (which is \textit{really} a lot of detections).
We now clearly state in Section \ref{M-sec:data}: ``This dataset includes 947 Ovenbirds counted in 381 surveys at 65 sites with site-specific variables including site age,  stock density, and an indicator of select-/partial-cut logging during the 1990s."

We believe this uncertainty accurately reflects the uncertainty in estimation
of detection probability and abundance. 
The apparent certainty in previously published studies relies heavily on the 
assumption of constant detection rate.
\end{quote}

@@

As the authors’ suggest, most point-count studies are designed to maximize availability, which is typically high for birds (e.g., 80-95\%) with perceptibility (e.g., the probability an observer detects a cue that is given by an animal) being the much larger source of bias.  Therefore, this model, without including a supplemental method to estimate perceptibility, is of limited use to many readers.
\begin{quote}
We agree.
This manuscript applies to removal-only analyses.
Still, the concept can be carried over to removal-distance analyses, as we describe in a paragraph of the Discussion.
The implementation will be more complicated and will vary depending on how removal- and distance-sampling are integrated.
\jarad{This is ongoing work for us.}
\end{quote}
Further, although the method may be more accurate even for constant detection data, the loss in precision is such that users will likely not opt to use it unless they suspect non-constant detection through time.  Further, given its presumably arduous computation times (albeit unknown), adding a perceptibility component to the model would likely greatly increase computational demands with unknown implications to bias and precision.  Including a paragraph in the Discussion that outlines when the model would be most beneficial to users (e.g., what type of data, surveys, or sample sizes) and explicitly stating drawbacks (e.g., computation time, loss of precision) would greatly improve this section.  
\begin{quote}
\adam{I do not deem this necessary.  We have not explicitly studied sample size.  The rest is addressed at various locations in the manuscript but does not, I think, require a dedication paragraph in the Discussion.}
\jarad{I'll have to re-read, but overall providing a summary for the reader
seems like a good idea.}
\end{quote}

@@

Second, although the original number of simulations was inadequate for characterizing model accuracy, it nonetheless highlighted scenarios of gross inaccuracy, such as: (i) omitting a mixture component in modeling a peaked mixture dataset, or (ii) assuming constant detection when the rate varies during the survey period.
\jarad{Is the second point necessary?}
\end{quote}

Second, the statistical properties of the estimator of p(det) are likely depending on the size of the true value of p(det). Therefore, I think Table 1 should have been repeated for a whole range of values of the true value of p(det).
(Is a boundary like the one in Olkin et al 1981, for your case, a hard one such that below it estimation is bad, and above it estimation is ``uniformly" good?).
Without variation in ``true" simulated scenarios, it is easy to inadvertendly ``stack the deck" in favor or against a particular combination of settings (Particularly since your ``true" value of p is high: 0.8). A wide range of simulated truths is also necessary because to be useful, these methods should be widely applicable in the
tropics (low N's, low p's very often) as well as in temperate forests (large N's, large p's), for example.
\begin{quote}
We have repeated the simulations in Sections \ref{M-sec:mixture} and \ref{M-sec:family} at true detection probabilities $\pdet = 0.50, 0.65, 0.80$, and 0.95.
The findings have significantly added to our results and conclusions.
\jarad{Should we summarize the results here?}

With reference to Olkin's finding that point estimates of $N_s$ are unstable unless $\pdet>0.41$, it is not a hard boundary.  
It varies by context, but the boundary is a general property of estimating $N$ in binomial unknown-$N$, unknown-$p$ problems.
Olkin's calculation derives from the relative sample mean and sample variance of repeated large samples with the same values of $N$ and $\pdet$.
As such, it is a best-case scenario.
Similar thresholds continue to recur in various abundance-estimating contexts, as demonstrated in these recent articles: 
\begin{itemize}
\item \citet{Veech2016}: advise caution when $p<0.5$ in the context of individual-level detection effects (no time-to-detection data). 
\item \citet{Field2016}: plots indicate uncertainty in $\hat{N}$ increase rapidly as $\pdet$ decreases below 0.40.  Methods include removal, double-observer, distance, and multiple-visit.
\item \citet{Davis2016}: report accurate $\hat{N}$ at $\pdet >0.40$ in a traditional removal-sampling context, though for small abundance ($N<50$) they required $\pdet > 0.7$.
\end{itemize}
\jarad{I don't think this parameter is necessary.}
\end{quote}

@@

We agree that discussions of identifiability and the flatness of the likelihood is very relevant, and we have introduced an explicit discussion of them in the Discussion.

As an informational note, the \texttt{R} package by Solymos was constructed for BUGS and JAGS, though it would not be difficult to adapt data cloning for Stan.
\jarad{Remove paragraph.}
\end{quote}

@@

Writing: The multiplicity of mathematical explicit meanings for the word ``mixture" may result confusing. The distinction between the mixture component a la Farnsworth et al and the N-mixture models (e.g. binomial counts with N~pois and p random) is clear to me, but it may not be for some of JABES' audience.  I suggest re-shaping the introduction to that effect. 
\begin{quote}
In the Introduction, we now clearly refer to a practice of modeling the TTDD ``as a mixture of two distributions --- a continuous-time distribution and a point mass for increased detection probability in the initial observation period."
We define our use of `mixture model' later in the paragraph:  ``We apply the term `mixture model' to any model with a mixture TTDD."
This is contrasted in our definition of `N-mixture model' in the next paragraph: ``a hierarchical framework for multinomial counts called an N-mixture model [\dotso] --- which is an entirely different use of `mixture' from the mixture models in the previous paragraph."
\jarad{We might not need the ``which..''.}
\end{quote}

@@

The authors could have crafted a beautiful and very telling graph (Figure 1) representing the multinomial intervals, the TTDD, the right censoring due to the end of the observation period, and the increased detection probability mixture component. Such figure (which is what I did by hand as I was reading the manuscript) would be a very useful guide in the reading of the paper.
\begin{quote}
We have added this as Figure \ref{M-fig:schematic}.
\jarad{Thank the reviewer for the suggestion.}
\end{quote}
